---
title: "STOPPD participant count by group "
output: html_document
---

written: January 18, 2018  
last run: January 25, 2018  

This script identifies the number of participants in olanzapine vs. placebo by scan timepoint, using the logic of group information that Judy and Dielle provided.    
__Note:__ this script includes data from all participants with data in Judy's master log and our file system. It has not excluded participants on any other basis (e.g., pilot data, ineligibility, etc.)


```{r setup, include=FALSE}

library('stringi')
library('plyr')
library('tidyr')
library('stringr')

df <- read.csv('../generated_csvs/STOPPD_masterDF_2018-01-18.csv') #clean spreadsheet created by nc
df[ df == "" ] <- NA #turn blank space into proper NAs

```

#First - identify the number of baseline scans (ie week 20) (as I understand it should be about 80).

```{r FirstScan}

#count the number of participants that have a 'yes' for 'completed' in "Scan.completed.1"
sum(na.omit(df$first_complete == "Yes")) #88 completed week 20 scan

#RANDOMIZATION 
(R <- addmargins(table(df$first_complete == 'Yes', df$randomization))) #O = 45; P = 43 (total = 88)

```

We see that 88 participants completed the week 20 scan, O = 45; P = 43 

#Second - identify the number of week 56 scans. (Ie 36 weeks after week 20).

```{r SecondScan_RCT}

#make sure that all the participants that completed week 56 scan also completed week 20
(df$second_complete == "Yes") %in% (df$first_complete== "Yes") #all TRUE

#count the number of participants that have a 'yes' for 'completed' in "Scan.completed" - but this includes 'relapse' and 'off protocol', as well as true completers
sum(na.omit(df$second_complete == "Yes")) #74 completed week 56 scan

#count how many participants that completed week 56 scan are classified as RCT (ignore 'relapse' and 'off protocol' for now)
sum(na.omit(df$second_complete == 'Yes' & df$second_timepoint == 'RCT')) #RCT = 41

#RANDOMIZATION 
(R <- addmargins(table(df$second_complete == 'Yes' & df$second_timepoint == 'RCT', df$randomization))) #O = 27; P = 14 (total = 41)


```

We see that 41 participants completed the week 56 scan, O = 27; P = 14


#Third - identify the number of "off label" scans also at week 56 (should only be a small handful).

```{r ThirdScan_OffLabel}

#make sure timepoint is a character
df$second_timepoint <- as.character(df$second_timepoint)

#count number of scans completed at *third* timepoint, which are by definition "off label"
sum(na.omit(df$third_complete == 'Yes')) #8 off-label scans

#of these, determine how many "off protocol" vs. "relapse", based on second timepoint scan
sum(na.omit(df$third_complete == 'Yes' & df$second_timepoint  == 'Off protocol')) #2 "off protocol" scans
sum(na.omit(df$third_complete == 'Yes' & df$second_timepoint == 'Relapse')) #6 relapse scans

#RANDOMIZATION 
(R <- addmargins(table(df$third_complete == 'Yes', df$randomization))) #Off label: O = 3; P = 5 (total = 8)
(R <- addmargins(table(df$third_complete == 'Yes' & df$second_timepoint == 'Off protocol', df$randomization))) #Off label - off protocol: O = 2; P = 0 (total = 2)
(R <- addmargins(table(df$third_complete == 'Yes' & df$second_timepoint == 'Relapse', df$randomization))) #Off label - relapse: O = 1; P = 5 (total = 6)

```

We see that 8 participants completed "off label" scans at week 56 scan, O = 3; P = 5.  Of these, 2 were "off protocol", O = 2; P = 0, and 6 were "relapse", O = 1; P = 5


#Then using our master log identify scans completed between week 20 and week 56 which are the relapse scans (and in a small minority of cases may be a scan when somebody is moving or wants out of the study despite being well). 

```{r SecondScan_Relapse}

#count relapse - note: both 'relapse' and 'off protocol' is included here (everything other than 'RCT')
sum(na.omit((df$second_timepoint == 'Relapse' | df$second_timepoint == 'Off protocol') & df$second_complete == 'Yes')) #33 participants relapsed/off protocol

#of these, count how many were "relapse" and how many were "off protocol"
sum(na.omit(df$second_timepoint == 'Relapse' & df$second_complete == 'Yes'))# 28 relapse
sum(na.omit(df$second_timepoint == 'Off protocol' & df$second_complete == 'Yes'))#5 off protocol

#RANDOMIZATION 
(R <- addmargins(table((df$second_timepoint == 'Relapse' | df$second_timepoint == 'Off protocol') & df$second_complete == 'Yes', df$randomization))) #relapse & off-protocol : O = 12; P = 21 (total = 33)
(R <- addmargins(table(df$second_timepoint == 'Relapse' & df$second_complete == 'Yes', df$randomization))) #relapse: O = 8; P = 20 (total = 28)
(R <- addmargins(table(df$second_timepoint == 'Off protocol' & df$second_complete == 'Yes', df$randomization))) #relapse: O = 4; P = 1 (total = 5)

```

We see that 33 participants completed a relapse scan, O = 12; P = 21. Of these, 28 were "relapse", O = 8; P = 20, and 5 were "off protocol", O = 4; P = 1.

