---
title: 'STOPPD participant count by group '
written: January 18, 2018  
last updated: August 20, 2018
last run: Sys.Date()
output:
  pdf_document: default
  html_document: default
---

This script identifies the number of participants in olanzapine vs. placebo by scan timepoint, using the logic of group inclusion that Judy and Dielle provided, and Nick and Aristotle have agreed to.

__Note:__ this script includes data from all participants with data in Judy's master log and our file system. It has not excluded participants on any other basis (e.g., QC fail, processing fail, post-hoc clinical trial ineligibility, etc.)


```{r setup, include=FALSE}

library('stringi')
library('plyr')
library('tidyr')
library('stringr')

df <- read.csv('../generated_csvs/STOPPD_masterDF_2018-09-06.csv', na.strings = '', stringsAsFactors = FALSE) #created by 03_STOPPD_masterDF

```

#First - identify the number of baseline scans (i.e., scans completed at week 20).

```{r FirstScan}

#count the number of participants that have a 'yes' for 'completed' in "Scan.completed.1"
sum(na.omit(df$first_complete == "Yes")) #88 participants completed week 20 scan

#for clarity, print the IDs of the N=88 participants that completed week 20 scans
(as.vector(na.omit(df$STUDYID[df$first_complete == "Yes"])))

#RANDOMIZATION - as expected, there's no difference in first scan completion between those randomized to O vs. P group
(R <- addmargins(table(df$first_complete == 'Yes', df$randomization))) #O = 45; P = 43 (total = 88)

```

#Second - identify the number of week 56 scans (i.e., 36 weeks after week 20).

```{r SecondScan_RCT}

#make sure that all the participants that completed week 56 scan also completed week 20
(df$second_complete == "Yes") %in% (df$first_complete== "Yes") #all TRUE

#count the number of participants that have a 'yes' for 'completed' in "Scan.completed" - but this includes 'relapse' and 'off protocol', as well as RCT or "true completers"
sum(na.omit(df$second_complete == "Yes")) #74 completed week 56 scan

#for clarity, print the IDs of the N=74 participants that completed week 56 scans
(as.vector(na.omit(df$STUDYID[df$second_complete == "Yes"])))

#count how many participants that completed week 56 scan are classified as RCT 
sum(na.omit(df$second_complete == 'Yes' & df$second_timepoint == 'RCT')) #RCT = 41
  (as.vector(na.omit(df$STUDYID[df$second_complete == "Yes" & df$second_timepoint == 'RCT']))) #for clarity, print N=41 RCT participant IDs
sum(na.omit(df$second_complete == 'Yes' & df$second_timepoint == 'Relapse')) #Relapse = 28
  (as.vector(na.omit(df$STUDYID[df$second_complete == "Yes" & df$second_timepoint == 'Relapse']))) #for clarity, print N=28 Relapse participant IDs
sum(na.omit(df$second_complete == 'Yes' & df$second_timepoint == 'Off protocol')) #Off protocol = 5
  (as.vector(na.omit(df$STUDYID[df$second_complete == "Yes" & df$second_timepoint == 'Off protocol']))) #for clarity, print N=5 Off protocol participant IDs

#RANDOMIZATION- look at randomization info for those who completed a second timepoint RCT scan
(R <- addmargins(table(df$second_complete == 'Yes' & df$second_timepoint == 'RCT', df$randomization))) #O = 27; P = 14 (total = 41)

```

#Third - identify the number of "off label" scans also at week 56.

```{r ThirdScan_OffLabel}

#make sure timepoint is a character
df$second_timepoint <- as.character(df$second_timepoint)

#count the number of scans completed at *third* timepoint, which are by definition "off label"
sum(na.omit(df$third_complete == 'Yes')) #8 off-label scans

#for clarity, print the IDs of the N=8 participants that completed off-label scans
(as.vector(na.omit(df$STUDYID[df$third_complete == "Yes"])))

#of these, determine how many "off protocol" vs. "relapse", based on second timepoint scan
sum(na.omit(df$third_complete == 'Yes' & df$second_timepoint  == 'Off protocol')) #2 "off protocol" scans
  (as.vector(na.omit(df$STUDYID[df$third_complete == "Yes" & df$second_timepoint  == 'Off protocol'])))
sum(na.omit(df$third_complete == 'Yes' & df$second_timepoint == 'Relapse')) #6 relapse scans
  (as.vector(na.omit(df$STUDYID[df$third_complete == "Yes" & df$second_timepoint  == 'Relapse'])))

#RANDOMIZATION 
(R <- addmargins(table(df$third_complete == 'Yes', df$randomization))) #Off label: O = 3; P = 5 (total = 8)
(R <- addmargins(table(df$third_complete == 'Yes' & df$second_timepoint == 'Off protocol', df$randomization))) #Off label - off protocol: O = 2; P = 0 (total = 2)
(R <- addmargins(table(df$third_complete == 'Yes' & df$second_timepoint == 'Relapse', df$randomization))) #Off label - relapse: O = 1; P = 5 (total = 6)

```

#Identify the scans completed between week 20 and week 56 which are the relapse scans (and in a small minority of cases may be a scan when somebody is moving or wants out of the study despite being well). 

```{r SecondScan_Relapse}

#count relapse - note: both 'relapse' and 'off protocol' is included here (everything other than 'RCT')
sum(na.omit((df$second_timepoint == 'Relapse' | df$second_timepoint == 'Off protocol') & df$second_complete == 'Yes')) #33 participants relapsed/off protocol

#of these, count how many were "relapse" and how many were "off protocol"
sum(na.omit(df$second_timepoint == 'Relapse' & df$second_complete == 'Yes'))# 28 relapse
sum(na.omit(df$second_timepoint == 'Off protocol' & df$second_complete == 'Yes'))#5 off protocol

#RANDOMIZATION 
(R <- addmargins(table((df$second_timepoint == 'Relapse' | df$second_timepoint == 'Off protocol') & df$second_complete == 'Yes', df$randomization))) #relapse & off-protocol : O = 12; P = 21 (total = 33)
(R <- addmargins(table(df$second_timepoint == 'Relapse' & df$second_complete == 'Yes', df$randomization))) #relapse: O = 8; P = 20 (total = 28)
(R <- addmargins(table(df$second_timepoint == 'Off protocol' & df$second_complete == 'Yes', df$randomization))) #relapse: O = 4; P = 1 (total = 5)

rm(df, R)

```
